---
title: "HMMtBroadPeak Guide"
author: "Jianhong Ou"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
package: "`r packageVersion('HMMtBroadPeak')`"
bibliography: bibliography.bib
csl: nature.csl
vignette: >
    %\VignetteIndexEntry{HMMtBroadPeak Vignette}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
output:
  html_document:
    theme: simplex
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r, echo=FALSE, results="hide", warning=FALSE}
suppressPackageStartupMessages({
  library(HMMtBroadPeak)
  library(rtracklayer)
  library(ggplot2)
})
knitr::opts_chunk$set(warning=FALSE, message=FALSE, eval=TRUE)
```

# Introduction

The HMMtBroadPeak package is designed to call very broad peaks for data
such as lamina-associated domains (LADs), 
nucleolus-associated domains (NADs), or other topologically associating 
domains.

The methods is following the description of Christ et.al[@LEEMANS2019852].
Reads will be count by each bins. Only bins with at least given reads
(defined by background parameter) for all samples (pool all reads for 
each bin) will be subsequently normalized. These bins will be first 
normalized to CPM (count per million) reads and then do log2 transform
for the ratio over control with a pseudocount. 
The peaks were defined by running a hidden markov model over the 
normalized values (using the R-package HMMt).

## Quick start

There are three steps for calling peaks:

### Step1: prepare the bam files.

The bam files should be clean with reads passed quality control and
proper paired (if applicable). The index file of bam should be stored
in the same folder and with same prefix.

```{r}
treatment <- system.file("extdata", "LB1.KD.chr1_1_5000000.bam",
                         package = "HMMtBroadPeak",
                         mustWork = TRUE)
control   <- system.file("extdata", "LB1.WT.chr1_1_5000000.bam",
                         package = "HMMtBroadPeak",
                         mustWork = TRUE)
## For local file, please try
# treatment <- "path/to/treatment/bam/files"
# control <- "path/to/control/bam/files"
```

### Step2: calling peaks.

The reads counts for treatment and control will be pool for each group. 
That is to say duplicates will not be considered when we call peaks.

```{r}
library(HMMtBroadPeak)
called <- HMMtBroadPeak(treatment, control)
called$peaks
```

### Step3: validate the calling and export peaks

```{r}
library(ggplot2)
plotPeaks(called, seqname="chr1") + theme_bw()
library(rtracklayer)
export(called$peaks, "called.broad.peak.bed")
```


# Session Info
```{r sessionInfo}
sessionInfo()
```

# References
